### 2021-11-30 placeholder.tex
Updates from Overleaf

\begin{thebibliography}{10}

\bibitem{proctor2019}
Proctor LM, Creasy HH, Fettweis JM, {Lloyd-Price} J, Mahurkar A, Zhou W, et~al.
\newblock The {{Integrative Human Microbiome Project}}.
\newblock Nature. 2019;569(7758):641--648.
\newblock doi:{10.1038/s41586-019-1238-8}.

\bibitem{sharma2019}
Sharma S, Tripathi P.
\newblock Gut Microbiome and Type 2 Diabetes: Where We Are and Where to Go?
\newblock The Journal of Nutritional Biochemistry. 2019;63:101--108.
\newblock doi:{10.1016/j.jnutbio.2018.10.003}.

\bibitem{aoun2020}
Aoun A, Darwish F, Hamod N.
\newblock The {{Influence}} of the {{Gut Microbiome}} on {{Obesity}} in
  {{Adults}} and the {{Role}} of {{Probiotics}}, {{Prebiotics}}, and
  {{Synbiotics}} for {{Weight Loss}}.
\newblock Preventive Nutrition and Food Science. 2020;25(2):113--123.
\newblock doi:{10.3746/pnf.2020.25.2.113}.

\bibitem{cho2012}
Cho I, Blaser MJ.
\newblock The Human Microbiome: At the Interface of Health and Disease.
\newblock Nature Reviews Genetics. 2012;13(4):260--270.
\newblock doi:{10.1038/nrg3182}.

\bibitem{callahan2016}
Callahan BJ, McMurdie PJ, Rosen MJ, Han AW, Johnson AJA, Holmes SP.
\newblock {{DADA2}}: {{High}}-Resolution Sample Inference from {{Illumina}}
  Amplicon Data.
\newblock Nature Methods. 2016;13(7):581--583.
\newblock doi:{10.1038/nmeth.3869}.

\bibitem{truong2015}
Truong DT, Franzosa EA, Tickle TL, Scholz M, Weingart G, Pasolli E, et~al.
\newblock {{MetaPhlAn2}} for Enhanced Metagenomic Taxonomic Profiling.
\newblock Nature Methods. 2015;12(10):902--903.
\newblock doi:{10.1038/nmeth.3589}.

\bibitem{li2019}
Li H.
\newblock Statistical and {{Computational Methods}} in {{Microbiome}} and
  {{Metagenomics}}.
\newblock In: Handbook of {{Statistical Genomics}}. {John Wiley \& Sons, Ltd};
  2019. p. 977--550.

\bibitem{li2015}
Li H.
\newblock Microbiome, {{Metagenomics}}, and {{High}}-{{Dimensional
  Compositional Data Analysis}}.
\newblock Annual Review of Statistics and Its Application. 2015;2(1):73--94.
\newblock doi:{10.1146/annurev-statistics-010814-020351}.

\bibitem{shi2016}
Shi P, Zhang A, Li H.
\newblock Regression Analysis for Microbiome Compositional Data.
\newblock The Annals of Applied Statistics. 2016;10(2):1019--1040.
\newblock doi:{10.1214/16-AOAS928}.

\bibitem{sankaran2014}
Sankaran K, Holmes S.
\newblock {{structSSI}}: {{Simultaneous}} and {{Selective Inference}} for
  {{Grouped}} or {{Hierarchically Structured Data}}.
\newblock Journal of statistical software. 2014;59(13):1--21.
\newblock doi:{10.18637/jss.v059.i13}.

\bibitem{benjamini1995}
Benjamini Y, Hochberg Y.
\newblock Controlling the {{False Discovery Rate}}: {{A Practical}} and
  {{Powerful Approach}} to {{Multiple Testing}}.
\newblock Journal of the Royal Statistical Society Series B (Methodological).
  1995;57(1):289--300.

\bibitem{chen2018}
Chen J, King E, Deek R, Wei Z, Yu Y, Grill D, et~al.
\newblock An Omnibus Test for Differential Distribution Analysis of Microbiome
  Sequencing Data.
\newblock Bioinformatics. 2018;34(4):643--651.
\newblock doi:{10.1093/bioinformatics/btx650}.

\bibitem{weiss2017}
Weiss S, Xu ZZ, Peddada S, Amir A, Bittinger K, Gonzalez A, et~al.
\newblock Normalization and Microbial Differential Abundance Strategies Depend
  upon Data Characteristics.
\newblock Microbiome. 2017;5(1).
\newblock doi:{10.1186/s40168-017-0237-y}.

\bibitem{mcmurdie2014}
McMurdie PJ, Holmes S.
\newblock Waste {{Not}}, {{Want Not}}: {{Why Rarefying Microbiome Data Is
  Inadmissible}}.
\newblock PLOS Computational Biology. 2014;10(4):e1003531.
\newblock doi:{10.1371/journal.pcbi.1003531}.

\bibitem{love2014}
Love MI, Huber W, Anders S.
\newblock Moderated Estimation of Fold Change and Dispersion for {{RNA}}-Seq
  Data with {{DESeq2}}.
\newblock Genome Biology. 2014;15(12):550.
\newblock doi:{10.1186/s13059-014-0550-8}.

\bibitem{quinn2019}
Quinn TP, Erb I, Gloor G, Notredame C, Richardson MF, Crowley TM.
\newblock A Field Guide for the Compositional Analysis of Any-Omics Data.
\newblock GigaScience. 2019;8(giz107).
\newblock doi:{10.1093/gigascience/giz107}.

\bibitem{quinn2018b}
Quinn TP, Erb I, Richardson MF, Crowley TM.
\newblock Understanding Sequencing Data as Compositions: An Outlook and Review.
\newblock Bioinformatics. 2018;34(16):2870--2878.
\newblock doi:{10.1093/bioinformatics/bty175}.

\bibitem{gloor2017}
Gloor GB, Macklaim JM, {Pawlowsky-Glahn} V, Egozcue JJ.
\newblock Microbiome {{Datasets Are Compositional}}: {{And This Is Not
  Optional}}.
\newblock Frontiers in Microbiology. 2017;8.
\newblock doi:{10.3389/fmicb.2017.02224}.

\bibitem{aitchison1999}
Aitchison J.
\newblock A {{Concise Guide}} to {{Compositional Data Analysis}}. 1999; p. 134.

\bibitem{kurtz2015}
Kurtz ZD, M{\"u}ller CL, Miraldi ER, Littman DR, Blaser MJ, Bonneau RA.
\newblock Sparse and {{Compositionally Robust Inference}} of {{Microbial
  Ecological Networks}}.
\newblock PLOS Computational Biology. 2015;11(5):e1004226.
\newblock doi:{10.1371/journal.pcbi.1004226}.

\bibitem{kaul2017}
Kaul A, Davidov O, Peddada SD.
\newblock Structural Zeros in High-Dimensional Data with Applications to
  Microbiome Studies.
\newblock Biostatistics. 2017;18(3):422--433.
\newblock doi:{10.1093/biostatistics/kxw053}.

\bibitem{kaul2017a}
Kaul A, Mandal S, Davidov O, Peddada SD.
\newblock Analysis of {{Microbiome Data}} in the {{Presence}} of {{Excess
  Zeros}}.
\newblock Frontiers in Microbiology. 2017;8.
\newblock doi:{10.3389/fmicb.2017.02114}.

\bibitem{khatri2012}
Khatri P, Sirota M, Butte AJ.
\newblock Ten {{Years}} of {{Pathway Analysis}}: {{Current Approaches}} and
  {{Outstanding Challenges}}.
\newblock PLOS Computational Biology. 2012;8(2):e1002375.
\newblock doi:{10.1371/journal.pcbi.1002375}.

\bibitem{goeman2007}
Goeman JJ, B{\"u}hlmann P.
\newblock Analyzing Gene Expression Data in Terms of Gene Sets: Methodological
  Issues.
\newblock Bioinformatics. 2007;23(8):980--987.
\newblock doi:{10.1093/bioinformatics/btm051}.

\bibitem{subramanian2005}
Subramanian A, Tamayo P, Mootha VK, Mukherjee S, Ebert BL, Gillette MA, et~al.
\newblock Gene Set Enrichment Analysis: {{A}} Knowledge-Based Approach for
  Interpreting Genome-Wide Expression Profiles.
\newblock Proceedings of the National Academy of Sciences.
  2005;102(43):15545--15550.
\newblock doi:{10.1073/pnas.0506580102}.

\bibitem{ashburner2000}
Ashburner M, Ball CA, Blake JA, Botstein D, Butler H, Cherry JM, et~al.
\newblock Gene {{Ontology}}: Tool for the Unification of Biology.
\newblock Nature genetics. 2000;25(1):25--29.
\newblock doi:{10.1038/75556}.

\bibitem{hanzelmann2013}
H{\"a}nzelmann S, Castelo R, Guinney J.
\newblock {{GSVA}}: Gene Set Variation Analysis for Microarray and
  {{RNA}}-{{Seq}} Data.
\newblock BMC Bioinformatics. 2013;14(1):7.
\newblock doi:{10.1186/1471-2105-14-7}.

\bibitem{frost2020a}
Frost HR.
\newblock Variance-Adjusted {{Mahalanobis}} ({{VAM}}): A Fast and Accurate
  Method for Cell-Specific Gene Set Scoring.
\newblock Nucleic Acids Research. 2020;48(16):e94--e94.
\newblock doi:{10.1093/nar/gkaa582}.

\bibitem{mclaren2019}
McLaren MR, Willis AD, Callahan BJ.
\newblock Consistent and Correctable Bias in Metagenomic Sequencing
  Experiments.
\newblock eLife. 2019;8:e46923.
\newblock doi:{10.7554/eLife.46923}.

\bibitem{egozcue2005}
Egozcue JJ, {Pawlowsky-Glahn} V.
\newblock Groups of {{Parts}} and {{Their Balances}} in {{Compositional Data
  Analysis}}.
\newblock Mathematical Geology. 2005;37(7):795--828.
\newblock doi:{10.1007/s11004-005-7381-9}.

\bibitem{chong2020}
Chong J, Liu P, Zhou G, Xia J.
\newblock Using {{MicrobiomeAnalyst}} for Comprehensive Statistical,
  Functional, and Meta-Analysis of Microbiome Data.
\newblock Nature Protocols. 2020;15(3):799--821.
\newblock doi:{10.1038/s41596-019-0264-1}.

\bibitem{tian2005}
Tian L, Greenberg SA, Kong SW, Altschuler J, Kohane IS, Park PJ.
\newblock Discovering Statistically Significant Pathways in Expression
  Profiling Studies.
\newblock Proceedings of the National Academy of Sciences.
  2005;102(38):13544--13549.
\newblock doi:{10.1073/pnas.0506577102}.

\bibitem{egozcue2003}
Egozcue JJ, {Pawlowsky-Glahn} V, {Mateu-Figueras} G, {Barcelo-Vidal} C.
\newblock Isometric {{Logratio Transformations}} for {{Compositional Data
  Analysis}}.
\newblock Mathematical Geology. 2003; p.~22.

\bibitem{silverman2017}
Silverman JD, Washburne AD, Mukherjee S, David LA.
\newblock A Phylogenetic Transform Enhances Analysis of Compositional
  Microbiota Data.
\newblock eLife. 2017;6:e21887.
\newblock doi:{10.7554/eLife.21887}.

\bibitem{barbie2009}
Barbie DA, Tamayo P, Boehm JS, Kim SY, Moody SE, Dunn IF, et~al.
\newblock Systematic {{RNA}} Interference Reveals That Oncogenic
  {{{\emph{KRAS}}}}-Driven Cancers Require {{TBK1}}.
\newblock Nature. 2009;462(7269):108--112.
\newblock doi:{10.1038/nature08460}.

\bibitem{quast2013}
Quast C, Pruesse E, Yilmaz P, Gerken J, Schweer T, Yarza P, et~al.
\newblock The {{SILVA}} Ribosomal {{RNA}} Gene Database Project: Improved Data
  Processing and Web-Based Tools.
\newblock Nucleic Acids Research. 2013;41(D1):D590--D596.
\newblock doi:{10.1093/nar/gks1219}.

\bibitem{delignette-muller2015}
{Delignette-Muller} ML, Dutang C.
\newblock Fitdistrplus: {{An R}} Package for Fitting Distributions.
\newblock Journal of Statistical Software. 2015;64(4):1--34.

\bibitem{benaglia2009}
Benaglia T, Chauveau D, Hunter DR, Young D.
\newblock Mixtools: {{An R}} Package for Analyzing Finite Mixture Models.
\newblock Journal of Statistical Software. 2009;32(6):1--29.

\bibitem{washburne2017}
Washburne AD, Silverman JD, Leff JW, Bennett DJ, Darcy JL, Mukherjee S, et~al.
\newblock Phylogenetic Factorization of Compositional Data Yields Lineage-Level
  Associations in Microbiome Datasets.
\newblock PeerJ. 2017;5:e2969.
\newblock doi:{10.7717/peerj.2969}.

\bibitem{morton2017}
Morton JT, Sanders J, Quinn RA, McDonald D, Gonzalez A, {V{\'a}zquez-Baeza} Y,
  et~al.
\newblock Balance {{Trees Reveal Microbial Niche Differentiation}}.
\newblock mSystems. 2017;2(1).
\newblock doi:{10.1128/mSystems.00162-16}.

\bibitem{efron2004}
Efron B.
\newblock Large-{{Scale Simultaneous Hypothesis Testing}}.
\newblock Journal of the American Statistical Association.
  2004;99(465):96--104.
\newblock doi:{10.1198/016214504000000089}.

\bibitem{wu2012}
Wu D, Smyth GK.
\newblock Camera: A Competitive Gene Set Test Accounting for Inter-Gene
  Correlation.
\newblock Nucleic Acids Research. 2012;40(17):e133.
\newblock doi:{10.1093/nar/gks461}.

\bibitem{calgaro2020}
Calgaro M, Romualdi C, Waldron L, Risso D, Vitulo N.
\newblock Assessment of Statistical Methods from Single Cell, Bulk {{RNA}}-Seq,
  and Metagenomics Applied to Microbiome Data.
\newblock Genome Biology. 2020;21(1):191.
\newblock doi:{10.1186/s13059-020-02104-1}.

\bibitem{cario1997}
Cario MC.
\newblock Modeling and {{Generating Random Vectors}} with {{Arbitrary Marginal
  Distributions}} and {{Correlation Matrix}}. 1997; p.~19.

\bibitem{agresti1998}
Agresti A, Coull BA.
\newblock Approximate {{Is Better}} than "{{Exact}}" for {{Interval
  Estimation}} of {{Binomial Proportions}}.
\newblock The American Statistician. 1998;52(2):119--126.
\newblock doi:{10.2307/2685469}.

\bibitem{delong1988}
DeLong ER, DeLong DM, {Clarke-Pearson} DL.
\newblock Comparing the Areas under Two or More Correlated Receiver Operating
  Characteristic Curves: A Nonparametric Approach.
\newblock Biometrics. 1988;44(3):837--845.

\bibitem{hawinkel2019}
Hawinkel S, Mattiello F, Bijnens L, Thas O.
\newblock A Broken Promise: Microbiome Differential Abundance Methods Do Not
  Control the False Discovery Rate.
\newblock Briefings in Bioinformatics. 2019;20(1):210--221.
\newblock doi:{10.1093/bib/bbx104}.

\bibitem{breiman2001}
Breiman L.
\newblock Random {{Forests}}.
\newblock Machine Learning. 2001;45(1):5--32.
\newblock doi:{10.1023/A:1010933404324}.

\bibitem{kuhn2020}
Kuhn M, Wickham H.
\newblock Tidymodels: {{Easily}} Install and Load the 'tidymodels' Packages;
  2020.

\bibitem{xiao2018}
Xiao J, Chen L, Yu Y, Zhang X, Chen J.
\newblock A {{Phylogeny}}-{{Regularized Sparse Regression Model}} for
  {{Predictive Modeling}} of {{Microbial Community Data}}.
\newblock Frontiers in Microbiology. 2018;9.
\newblock doi:{10.3389/fmicb.2018.03112}.

\bibitem{dong2020}
Dong M, Li L, Chen M, Kusalik A, Xu W.
\newblock Predictive Analysis Methods for Human Microbiome Data with
  Application to {{Parkinson}}'s Disease.
\newblock PLOS ONE. 2020;15(8):e0237779.
\newblock doi:{10.1371/journal.pone.0237779}.

\bibitem{pasolli2017}
Pasolli E, Schiffer L, Manghi P, Renson A, Obenchain V, Truong DT, et~al.
\newblock Accessible, Curated Metagenomic Data through {{ExperimentHub}}.
\newblock Nature Methods. 2017;14(11):1023--1024.
\newblock doi:{10.1038/nmeth.4468}.

\bibitem{schiffer2019}
Schiffer L, Azhar R, Shepherd L, Ramos M, Geistlinger L, Huttenhower C, et~al.
\newblock {{HMP16SData}}: {{Efficient}} Access to the Human Microbiome Project
  through Bioconductor.
\newblock American Journal of Epidemiology. 2019;doi:{10.1093/aje/kwz006}.

\bibitem{gonzalez2018}
Gonzalez A, {Navas-Molina} JA, Kosciolek T, McDonald D, {V{\'a}zquez-Baeza} Y,
  Ackermann G, et~al.
\newblock Qiita: Rapid, Web-Enabled Microbiome Meta-Analysis.
\newblock Nature Methods. 2018;15(10):796--798.
\newblock doi:{10.1038/s41592-018-0141-9}.

\bibitem{consortium2012}
Consortium THMP, Huttenhower C, Gevers D, Knight R, Abubucker S, Badger JH,
  et~al.
\newblock Structure, Function and Diversity of the Healthy Human Microbiome.
\newblock Nature. 2012;486(7402):207--214.
\newblock doi:{10.1038/nature11234}.

\bibitem{thurnheer2016}
Thurnheer T, Bostanci N, Belibasakis GN.
\newblock Microbial Dynamics during Conversion from Supragingival to
  Subgingival Biofilms in an in Vitro Model.
\newblock Molecular Oral Microbiology. 2016;31(2):125--135.
\newblock doi:{10.1111/omi.12108}.

\bibitem{beghini2019}
Beghini F, Renson A, Zolnik CP, Geistlinger L, Usyk M, Moody TU, et~al.
\newblock Tobacco Exposure Associated with Oral Microbiota Oxygen Utilization
  in the {{New York City Health}} and {{Nutrition Examination Study}}.
\newblock Annals of Epidemiology. 2019;34:18--25.e3.
\newblock doi:{10.1016/j.annepidem.2019.03.005}.

\bibitem{matteocalgaro2020}
Calgaro M. Mcalgaro93/Sc2meta: {{Paper Release}}; 2020.
\newblock Zenodo.

\bibitem{nielsen2014}
Nielsen HB, Almeida M, Juncker AS, Rasmussen S, Li J, Sunagawa S, et~al.
\newblock Identification and Assembly of Genomes and Genetic Elements in
  Complex Metagenomic Samples without Using Reference Genomes.
\newblock Nature Biotechnology. 2014;32(8):822--828.
\newblock doi:{10.1038/nbt.2939}.

\bibitem{gevers2014}
Gevers D, Kugathasan S, Denson LA, {V{\'a}zquez-Baeza} Y, Van~Treuren W, Ren B,
  et~al.
\newblock The {{Treatment}}-{{Naive Microbiome}} in {{New}}-{{Onset Crohn}}'s
  {{Disease}}.
\newblock Cell Host \& Microbe. 2014;15(3):382--392.
\newblock doi:{10.1016/j.chom.2014.02.005}.

\bibitem{martin2020}
Martin BD, Witten D, Willis AD.
\newblock Modeling Microbial Abundances and Dysbiosis with Beta-Binomial
  Regression.
\newblock The Annals of Applied Statistics. 2020;14(1):94--115.
\newblock doi:{10.1214/19-AOAS1283}.

\bibitem{naim2012}
Naim I, Gildea D.
\newblock Convergence of the {{EM Algorithm}} for {{Gaussian Mixtures}} with
  {{Unbalanced Mixing Coefficients}}.
\newblock Proceedings of the 29th International Coference on International
  Conference on Machine Learning. 2012; p.~8.
\newblock doi:{10.5555/3042573.3042756}.

\bibitem{wu2016}
Wu C, Chen J, Kim J, Pan W.
\newblock An Adaptive Association Test for Microbiome Data.
\newblock Genome Medicine. 2016;8(1):56.
\newblock doi:{10.1186/s13073-016-0302-3}.

\bibitem{martin-fernandez2012}
{Mart{\'i}n-Fern{\'a}ndez} JA, Hron K, Templ M, Filzmoser P,
  {Palarea-Albaladejo} J.
\newblock Model-Based Replacement of Rounded Zeros in Compositional Data:
  {{Classical}} and Robust Approaches.
\newblock Computational Statistics \& Data Analysis. 2012;56(9):2688--2704.
\newblock doi:{10.1016/j.csda.2012.02.012}.

\end{thebibliography}