### 2022-01-27 supplemental.tex

\documentclass{article}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{float}
\usepackage[parfill]{parskip}
\newcommand{\overbar}[1]{\mkern 2mu\overline{\mkern-2mu#1\mkern-20mu}\mkern 20mu}

\title{Supplementary Materials for CBEA: Competitive Balances for Taxonomic Enrichment Analysis}
\begin{document}
\author{Quang P. Nguyen, Anne G. Hoen, H. Robert Frost}
\maketitle
\captionsetup[figure]{labelfont={bf},name={Figure},labelsep=period, margin=1cm}

\section{Addressing variance inflation due to correlation}

In order to perform inference with CBEA, we estimated the null distribution empirically. This can be done either non-parametrically by constructing a null distribution through computing scores on multiple permutations of the data, or parametrically via estimating parameters of a distribution using the same permuted scores. This means that our null distribution for a given set is equivalent to scores computed for sets of similar sizes but containing randomly chosen taxa. We chose two distributional forms for the null: the normal distribution and a two-component mixture normal distribution. For the normal distribution, we estimated parameters using the maximum likelihood using the \emph{fitdistrplus} package \cite{delignette-muller2015}. For the mixture normal distribution, we utilized the expectation-maximization procedure from the package \emph{mixtools} \cite{benaglia2009}.

An advantage of estimating the null using a parametric approach is the ability to correct for variance inflation due to inter-taxa correlation within the set compared to overall background correlation \cite{wu2012}. CBEA addresses this issue by combining the location (or mean) estimate from the column permuted raw score matrix with the spread (or variance) estimate from the original un-permuted scores. This still allows us to leverage the null generated via column permutation while using the proper variance estimate taken from scores where the correlation structure has not been disrupted. As such, this procedure assumes that the variance of the test statistic under the alternate hypothesis is the same as that of the null.

For the normal distribution, it is straightforward to combine mean and variance estimates from the respective raw score matrices. For the mixture normal distribution, however, due to the fact that the distribution is made of component-wise means, variances and mixing coefficients, we decided to take an optimization based approach to identifying the component wise variances $\sigma_1$ and $\sigma_2$ such the overall mean and variance estimates come from the respective raw cILR score matrices as detailed above. We can write the optimization problem as follows:

\begin{equation} \label{eq:3}
\begin{aligned}
    \min_{\sigma_1, \sigma_2} \quad & \sqrt{\left(\sqrt{(\sigma_1 + \mu'_1 - M')\lambda'_1 + (\sigma_2 + \mu'_2 - M')\lambda'_2} - SD\right)^2}\\
    \textrm{s.t.}  \quad & \sigma_1 \geq 10^{-5}, \sigma_2 \geq 10^{-5}    \\
    \textrm{where} \quad & M', SD, \mu'_1, \mu'_2, \lambda'_1, \lambda'_2 \text{ are constants} \\
\end{aligned}
\end{equation}

$\lambda'_1$, $\lambda'_2$, $\mu'_1$, $\mu'_2$, and $M'$ are component-wise mixing coefficients, component-wise means, and overall mean of the mixture distribution estimated from column-permuted scores while $SD$ is the overall standard deviation of the mixture distribution estimated from unpermuted scores. We solve for this optimization problem using a quasi Newton method with box constraints (L-BFGS-B) with the default finite-difference approximation of the gradient, as implemented in the \emph{optim} function in R.

There are also different variations to this approach. We can choose to vary both $\sigma_1$ and $\sigma_2$ or to keep either $\sigma_1$ or $\sigma_2$ constant and varying the remaining component. We can assume that null distribution is a two-component mixture distribution where there is one major component with smaller mean (representing the bulk of the distribution centered at the permuted mean) and one minor component with higher mean (representing the inflated right tail). Under this assumption, we can modify the optimization problem to only estimate the variance parameter of the smaller component (i.e. withouth loss of generalizability keeping $\sigma_1$ constant where $\lambda'_1 > \lambda'_2$). This allows for the optimization procedure to more properly capture the right tail distribution rather than increasing weight on the left tail of the distribution which impacts the computation of p-values for a one-sided test. Empirical experiments (data not shown) done on simulated data and random set analyses suggest that this adjustment improves performance of the adjusted CBEA under mixture-normal assumption. In the R implementation of CBEA, users can control this behaviour by specifying the

\section{Performance evaluation} 
We implemented cILR in the package \emph{teaR} on GitHub with \emph{Rcpp} implementations for geometric mean computations. We evaluated computational time using the \emph{bench} package in R. Sample size and number of sets were varied across different evaluation conditions. All set sizes were kept constant at 50.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=\textwidth]{figures/performance.png}
    \caption{Computational time (in seconds) as a function of sample size (left panel) and number of taxa sets evaluated (right panel). Evaluation was performed on simulated data sets. For sample size analysis, only 10 sets were evaluated. For taxa set analysis, sample size was fixed at 1,000. Across all evaluations, the size of each taxa set was also fixed at 50.}
    \label{fig:s1}
\end{figure}

In general, increasing sample size and number of evaluated sets both increase computational time. The mixture normal distribution takes longer to compute compared to the normal distribution, and adjusted methods take longer but was most apparent when combined with the mixture distribution. Overall, a reasonable data set of 1,000 samples and 100 sets (5,000 taxa) would take around 200 seconds (or approximately 3 minutes) for the longest approach (correlation adjusted mixture normal distribution). Benchmark was performed on a single core using a personal Windows laptop (Specifications: Intel Core i7-9750H (2.6-4.5 GHz - 12MB Cache) with 16GB of RAM)

\section{Distribution of p-values}
We also demonstrate the validity our inference procedure by assessing the distribution of p-values. We used simulated data sets as described in the main manuscript, under different data sparsity levels (0, 0.6) and inter-taxa correlation within the set ($\rho$ = 0, 0.5). We evaluated p-values of the correlation adjusted method for $\rho = 0.5$ case and the p-values of the unadjusted method for $\rho = 0$ case.

In general, sparsity has no effect on the distribution of p-values, where it remains uniform at zero correlation values. However, for higher correlation values, p-values deviate more from uniform under the global null, with the mixture normal distribution being further away from uniform compared to the normal distribution.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=\textwidth]{figures/pval_distr.png}
    \caption{Q-Q plot of 10,000 p-values compared against a uniform distribution bounded between 0 and 1. Evaluation was performed on simulated null data sets of 10,000 samples testing for enrichment of a set of size 50. For correlation of 0.5, p-values represent correlation adjusted cILR while for correlation of 0, p-values represent unadjusted cILR.}
    \label{fig:s2}
\end{figure}

\newpage
\bibliography{mac_taxagg}{}
\bibliographystyle{plos2015}

\end{document}